Performance benefits of Intel\textsuperscript{\textregistered}~Optane\texttrademark~DC persistent memory for the parallel processing of large neuroimaging data

### 2019-12-17 paper.tex
tristan modifs

\begin{abstract}
    Open-access neuroimaging datasets have reached petabyte scale, and continue to grow.
The ability to leverage the entirety of these datasets is limited to a restricted
number of labs with both the capacity and infrastructure to
process the data. Whereas Big Data engines have significantly
reduced application performance penalties with respect to data
movement, their applied strategies are not necessarily practical
within neuroimaging workflows where intermediary results may
need to be materialized to shared storage for post-processing
analysis. In this paper we evaluate the performance advantage
brought by Intel\textsuperscript{textregistered}~Optane\texttrademark~DC persistent memory for the
processing of large neuroimaging datasets using the two available
configurations modes: Memory Mode and App Direct Mode. We
employ a synthetic algorithm on the 76 GiB and 603 GiB BigBrain, as well as apply a
standard neuroimaging application on the Consortium for Reliability
and Reproducibility (CoRR) dataset using 25 and 96 parallel
processes in both cases. Our results show that the performance of
the applications leveraging persistent memory is superior to that of other
storage devices, with the exception of DRAM. This is the case
in both Memory and App Direct Mode and irrespective of the
amount of data and parallelism. Furthermore, persistent memory in App
Direct Mode is believed to benefit from the use of memory as
a cache for writing in cases where output data is significantly
smaller than available memory. We believe the use of persistent memory
will be beneficial to both neuroimaging applications running on
HPC or visualization of large, high-resolution images.
\end{abstract}

\section{Introduction}
Neuroimaging open-data initiatives have led to extensively large repositories of
publicly available data. Such initiatives include the \bigbrain~\cite{BigBrain}, 
a one-of-a-kind $\SI{603}{\gibi\byte}$
histological image of a 65 year-old healthy human brain
at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
contain approximately $\SI{0.2}{\peta\byte}$ of data (including various magnetic
resonance (MR) imaging modalities) from 500,000 individuals living in the UK;
the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
1,200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
initiative which aggregates MR data from various centres around the world, 
of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
data in total.

Due to storage limitations, only subsets of such neuroimaging repositories 
can be processed in a typical research laboratory. Moreover, as these datasets 
are extremely large and are only increasing in size, they are typically stored 
in higher-capacity, slower storage devices, such as hard disk drives, or external
parallel file systems, such as Lustre. In such conditions, large-scale 
studies in neuroimaging remain limited to labs with access to adequate 
infrastructures.

@@

To mitigate the effects of data writes, kernel-based
strategies, such as the writeback cache, have been developed. The writeback cache
allows processes to use the memory as a cache for writing. The cache size,
however,
is configurable, but nevertheless limited. When writes approach the cache's capacity,
processes performing writes start to be throttled. Once the cache capacity is reached,
processes can no longer use the cache until all cached written data is flushed
to the appropriate storage device.

@@

Intel Optane DC Persistent Memory~\cite{optanebrief} is a high-performance
storage technology that resides on the DIMMs to reduce latency to the device.
The first generation of Intel Optane DC persistent memory offers capabities of 128~GB,
256~GB, and 512~GB, enabling it to be more cost effective than DRAM storage.
There are two configuration modes, Memory Mode and App Direct Mode,
that enable the storage to either be accessed as an extension of volatile main memory
or as a non-volatile memory storage device\footnote{https://itpeernetwork.intel.com/intel-optane-dc-persistent-memory-operating-modes/}.

In this paper, we aim to:
\begin{itemize}
        \item Quantify the added value of Intel Optane DC persistent memory on 
            processing large neuroimaging data using representative pipelines; and
        \item Determine when a given Intel Optane DC persistent memory configuration (Memory 
            and App Direct Mode) is preferable.
\end{itemize}

@@

The server used for the experiments consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\giga\byte} Intel Optane DC persistent memory
modules, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC persistent memory.
Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
only \SI{149}{\gibi\byte} where available for the experiments, and a
\SI{720}{\tebi\byte} shared Dell EMC Isilon cluster. The Isilon cluster is an nsf4 mounted
cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
was the only storage device set up as a writeback device. Both Isilon and Optane DC persistent memory in
App Direct mode were configured as write-through devices, meaning they did not leverage
DRAM as a cache for writes. For Optane DC persistent memory, this was achieved by configuring the xfs filesystem with
DAX. All storage devices were benchmarked using the following \href{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}{script}.
The result of the benchmarks can be seen in Table \ref{table:bandwidths}.

For processing,
2 Intel(R) Xeon(R) Platinum 8260M CPU @ 2.40GHz where installed, enabling use of up to 96 threads.
The server was running Red Hat Enterprise Linux 7.6 release (Maipo).

@@

\subsubsection{Memory Mode}

@@

While enabling access to a larger amount of memory than typically accessible otherwise,
it is anticipated that Memory mode will be slower than App Direct mode for all memory accessed that
are not cached in DRAM. We evaluated all available devices (Optane DCPMM through tmpfs, local disk and Isilon)
in Memory mode.

\subsubsection{App Direct Mode}

App Direct mode enables Optane DCPMM to be accessed as a high-performance storage device.
Unlike Memory mode, the OS is able to differentiate between DRAM and Optane DCPMM,
treating them as two distinct memory tiers. Optane DCPMM
does not use the DRAM as cache in App Direct Mode if configured with DAX.

@@

We characterize the performance of our experiments using the following model:

@@

For applications which have negligible CPU time, as is the case with our experiments, 
it is expected that the I/O duration can estimate the total makespan. However, there
are certain instances where the makespan may be below the I/O duration. This is expected
to occur with storage devices that scale well with increased parallelism, as should be the
case for DRAM, Optane DCPMM and Isilon, which all consist of multiple devices. Should the
scalable storage devices predict the makespan accurately, it is believed that other overheads
would be at play making the device unfavourable for parallel processing.

@@

Due to the size and uniqueness of the \bigbrain, standardized processing pipelines
have yet to be developed. In order to quantify the effects of storage devices
on such a dataset, we have implemented a naive synthetic application that takes 
the image, split into blocks, and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
This application
enables us to read and write to different storage devices and ensure that written
data could not have been previously cache
in-memory, by ensuring that read and written data are not the same. This application
was parallelized in two different ways: using both GNU Parallel and Apache Spark 2.4.3 (PySpark). 
All code was implemented in Python 3.6.

There are some notable differences between the GNU Parallel and Apache Spark implementations.
For instance, in the GNU Parallel implementation, all operations (i.e. reading, incrementing and writing)
are done within a single task. This task is applied to all the \bigbrain blocks using GNU Parallel,
processing a subset of them at a time,
depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
is achieved using the popular neuroimaging I/O: \href{https://nipy.org/nibabel/}{NiBabel}. When provided with a filename, as is the case
for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memory map the data
using \href{https://numpy.org/}{NumPy}. This step will be referred to as ``load header''. It is only when the data is actually required
(i.e. during incrementation), that the data will be loaded into memory.

The Apache Spark implementation differs from that of GNU Parallel in that scheduling decisions
were entirely differed to Spark. That is, rather than simply parallelizing over the \bigbrain block filenames,
we opted to read the data using Spark's built-in
\texttt{BinaryFiles}, loading whole files, in binary format, into Spark partitions. As the
data would have been pre-loaded by Spark, ``loading header'' only measures the time to
convert the binary images into NiBabel objects, and ``incrementation'' consists solely
of the time it took to increment the data. Therefore, only the write time would be
comparable between the two implementations, as it is achieved using the same method.
Furthermore, the Spark implementation also differs in that read, increment and write
were separated into three map tasks, which can enable shuffling between the tasks should 
it be determined as necessary by the Spark scheduler.

@@

The BIDS App Example is a template example for creating a Brain Imaging Data Structure (BIDS)
compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
on all the anatomical (T1W) images of dataset containing numerous subjects. This step is 
referred to as Participant analysis within the application. An optional
step of the BIDS App example, referred to as Group analysis, computes the average brain
mask size of the entire dataset.

For our experiments, we used the entire CoRR dataset, made available on \href{}{DataLad} and
applied Participant analysis to it. The BIDS App Example was executed using a Singularity container stored on
Isilon. As in the BigBrain Incrementation, the experiment was parallelized 
using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode,
where Isilon and the local SSD were only evaluated in Memory Mode. Each experiment was repeated 3x.

Time in this application was obtained using Linux's \texttt{time(1)} application.
\texttt{Real + Sys} was used to measure CPU time,
whereas \texttt{User - (Real + Sys)} was used to measure
I/O time.
\section{Results}

\subsection{40~$\mu$m \bigbrain incrementation}
\subsubsection{25 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Makespan of the \bigbrain Incrementation application using 25 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. 3 
             repetitions were performed. *Local SSD did not complete the
    writing of the last few blocks due to storage limitations.}\label{fig:makespan-25cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_25cpus.pdf}
    \captionsetup{belowskip=0pt}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-25cpus}
\end{figure}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572380543-tmpfsAD_1it_25cpus_40bb-1.pdf}
    \caption{DRAM}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382482-optaneAD_1it_25cpus_40bb-1.pdf}
    \caption{Optane DCPMM}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382794-localAD_1it_25cpus_40bb-1.pdf}
    \caption{local SSD*}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572388118-isilonAD_1it_25cpus_40bb-2.pdf}
    \caption{Isilon}\label{fig:gantt25isilon}
\end{subfigure}
    \captionsetup{belowskip=-10pt}
\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes. *Some local SSD writes did not complete due to storage limitations.}\label{fig:gantt25}
\end{figure*}

@@

The App Direct mode Gantt charts (Figure~\ref{fig:gantt25}) also reflect what is observed in Figure~\ref{fig:stacked-25cpus}.
DRAM is significantly faster than all other storage, leading it to barely appear within the Gantt chart.
DRAM was measured to have an average of 17.2 parallel tasks throughout the execution. Optane DCPMM takes
significantly longer than DRAM and appears to spend a significant amount of time writing. The average
parallelism measured for Optane DCPMM was 24. Unlike Optane DCPMM, local SSD spends the vast majority of its time
reading, while spending a negligible amount of time writing. Like Optane DCPMM, it averages 24 parallel tasks
throughout its execution. Unlike the other three storage, Isilon shows spacing between the loading of the header,
read and increment and write. Isilon was measured to have an average parallelism of 14. Overall, each storage device
execution displayed that some tasks took longer than others.

\subsubsection{96 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_96cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Makespan of the \bigbrain Incrementation application using 96 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. 3 
             repetitions were performed. *Local SSD did not complete the
    writing of the last few blocks due to storage limitations.}\label{fig:makespan-96cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_96cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-96cpus}
\end{figure}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572306855-tmpfs_1it_96cpus_40bb-2.pdf}
    \caption{Memory Mode}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572379747-optaneAD_1it_96cpus_40bb-1.pdf}
    \caption{App Direct Mode}
\end{subfigure}
    \captionsetup{belowskip=-10pt}
\caption{Gantt charts for Optane DCPMM processing 125 blocks of the 40$\mu$m BigBrain using 96 processes}\label{fig:gantt96}
\end{figure*}

Despite having increased parallelism by a factor of 3.84, we see no visible reduction
in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 cpus, DRAM and Optane DCPMM in Memory mode are the only devices that did better than the anticipated I/O, with a makespan duration of
5.7s and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
Optane DCPMM in App Direct mode and local SSD both performed worse than the anticipated I/O estimates, Isilon
performed as expected. Variance was found to be high on the local SSD.

@@

A further look at the Gantt charts between Optane DCPMM in Memory mode and App Direct mode (Figure~\ref{fig:gantt96})
show that both read and write time of tasks are worse in App Direct mode as compare with Memory mode.
While some read tasks in App Direct mode appear to be of similar duration to the average read task time
of Memory mode, the vast majority of read tasks are significantly slower. Furthermore, no write tasks in App
Direct mode is capable of reaching the performance of equivalent tasks in Memory mode.

@@

Unlike the GNU Parallel executions, the Spark executions appear to be in better accordance with
the I/O estimates (Figure~\ref{fig:20mkspspark25}. Only Optane DCPMM in App Direct mode appears
to have significantly exceeded the estimates, by a factor of 2. Total write duration all appear
to be longer in the Spark implementation than with GNU Parallel (Figure~\ref{fig:20totalspark:25}).
Spark's writes are 1.4-1.6x longer with Optane DCPMM and 1.1x longer with Isilon. Due to differences in
how data is loaded in Spark, it is not possible to comment on how read times are affected. However, it
is possible to note that data conversion (binary string to NumPy array) takes almost twice as 
much time on Optane DCPMM than it does on Isilon. Furthermore, the act of incrementation can take
up to twice as much time on Isilon.

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_25cpus.pdf}
    \caption{Makespan}\label{fig:20mksp25}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
    \caption{Total read/increment/write breakdown}\label{fig:20total25}
\end{subfigure}
\captionsetup{belowskip=-10pt}
\caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
25 processes. 3 repetitions were performed.}\label{fig:2025}
\end{figure*}

>>>>

\begin{abstract}
    Open-access neuroimaging datasets have reached petabyte scale, and continue to grow.
The ability to leverage the entirety of these datasets is limited to a restricted
number of labs with both the capacity and infrastructure to
process the data. Whereas Big Data engines have significantly
reduced application performance penalties with respect to data
movement, their applied strategies (e.g. data locality, in-memory computing and lazy evaluation)
are not necessarily practical
within neuroimaging workflows where intermediary results may
need to be materialized to shared storage for post-processing
analysis. In this paper we evaluate the performance advantage
brought by Intel\textsuperscript{textregistered}~Optane\texttrademark~DC persistent memory for the
processing of large neuroimaging datasets using the two available
configurations modes: Memory mode and App Direct mode. We
employ a synthetic algorithm on the 76 GiB and 603 GiB BigBrain, as well as apply a
standard neuroimaging application on the Consortium for Reliability
and Reproducibility (CoRR) dataset using 25 and 96 parallel
processes in both cases. Our results show that the performance of
the applications leveraging persistent memory is superior to that of other
storage devices, with the exception of DRAM. This is the case
in both Memory and App Direct mode and irrespective of the
amount of data and parallelism. Furthermore, persistent memory in App
Direct mode is believed to benefit from the use of DRAM as
a cache for writing when output data is significantly
smaller than available memory. We believe the use of persistent memory
will be beneficial to both neuroimaging applications running on
HPC or visualization of large, high-resolution images.
\end{abstract}

\section{Introduction}
Neuroimaging open-data initiatives have led to extensively large repositories of
publicly available data. Such initiatives include the \bigbrain~\cite{BigBrain}, 
a one-of-a-kind $\SI{603}{\gibi\byte}$
histological image of a 65 year-old healthy human brain
at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
contain approximately $\SI{0.2}{\peta\byte}$ of data (including various magnetic
resonance (MR) imaging modalities) from 500,000 individuals living in the UK;
the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
1,200 healthy adults, expected to exceed $\SI{2}{\peta\byte}$ in size; 
and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
initiative which aggregates MR data from various centres around the world, 
of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
data in total.

Due to storage limitations, only subsets of such neuroimaging repositories 
can be processed in a typical research laboratory. Moreover, as these datasets 
are extremely large and are only increasing in size, they are typically stored 
in higher-capacity, slower storage devices, such as hard disk drives, or external
parallel file systems, such as Lustre~\cite{lustre}. In such conditions, large-scale 
studies in neuroimaging remain limited to labs with access to adequate 
infrastructures.

@@

To mitigate the effects of data writes, the Linux kernel has implemented
strategies, such as the writeback cache. The writeback cache
allows processes to use the memory as a cache for writing. This strategy contrasts writethrough, which 
writes data directly to the device. The writeback cache size
is configurable, but nevertheless limited. When writes approach the cache's capacity,
processes performing writes start to be throttled. Once the cache capacity is reached,
processes can no longer use the cache until all cached written data is flushed
to the appropriate storage device.

@@

Intel Optane DC Persistent Memory Module~\cite{optanebrief} (DCPMM) is a high-performance
storage technology that resides on the DIMMs to reduce latency to the device.
The first generation of Intel Optane DCPMM offers capabities of 128~GB,
256~GB, and 512~GB, enabling it to be more cost effective than DRAM storage.
There are two configuration modes, Memory mode and App Direct mode,
that enable the storage to either be accessed as an extension of volatile main memory
or as a non-volatile memory storage device\footnote{https://itpeernetwork.intel.com/intel-optane-dc-persistent-memory-operating-modes/}.

In this paper, we aim to:
\begin{itemize}
        \item Quantify the added value of Intel Optane DC persistent memory on 
            processing large neuroimaging data using representative pipelines; and
        \item Determine when a given Intel Optane DC persistent memory configuration (Memory 
            and App Direct mode) is preferable.
\end{itemize}

@@

The server used consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\giga\byte} Optane DCPMM
modules, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC persistent memory.
Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
only \SI{149}{\gibi\byte} where available for the experiments, and a
\SI{720}{\tebi\byte} shared Dell EMC Isilon network-attached storage platform. Isilon is an nsf4 mounted
cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
was the only storage device set up as a writeback device. Both Isilon and Optane DCPMM in
App Direct mode were configured as write-through devices, meaning they did not leverage
DRAM as a cache for writes. For Optane DC persistent memory, this was achieved by configuring the xfs filesystem with
direct access (DAX). DAX enables persistent memory to be byte addressable, bypassing the page cache. All storage devices were benchmarked using the following script: \url{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}.
The result of the benchmarks can be seen in Table \ref{table:bandwidths}.

For processing,
2 Intel(R) Xeon(R) Platinum 8260M CPU @ 2.40GHz where installed, enabling use of up to 96 threads.
The server was running Red Hat Enterprise Linux 7.6 (Maipo) kernel version 3.10.0-957.

@@

\subsubsection{Memory mode}

@@

While enabling access to a larger amount of memory than typically accessible otherwise,
it is anticipated that Memory mode will be slower than App Direct mode for all memory accesses that
are not cached in DRAM, due to reduced storage bandwidth. We evaluated all available devices (Optane DCPMM through tmpfs, local SSD and Isilon)
in Memory mode.

\subsubsection{App Direct mode}

App Direct mode enables Optane DCPMM to be accessed as a high-performance storage device.
Unlike Memory mode, the OS is able to differentiate between DRAM and Optane DCPMM,
treating them as two distinct memory tiers. Optane DCPMM
does not use the DRAM as cache in App Direct Mode when configured with DAX, as is our case.

@@

We characterize the performance of our data-intensive experiments using the following model:

@@

For applications which have negligible CPU time, as is the case with our experiments, 
it is expected that the I/O duration can estimate the total makespan. However, there
are certain instances where the makespan may be below the I/O duration. This is expected
to occur with scalable devices, as should be the
case for DRAM, Optane DCPMM and Isilon, which can support parallel I/O. Should the
scalable storage devices predict the makespan accurately, it is believed that other overheads
would be at play, making the device unfavourable for parallel processing.

@@

Due to the size and uniqueness of the \bigbrain, standardized processing pipelines
have yet to be developed. In order to quantify the effects of storage devices
on such a dataset, we have implemented a naive synthetic application that takes 
the image, split into blocks, and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
This application
enables us to read and write to different storage devices and ensure that written
data could not have been previously cached
in-memory, by ensuring that read and written data are not the same. This application
was parallelized in two different ways: using both GNU Parallel~\cite{gnuparallel} and Apache Spark 2.4.3 (PySpark)~\cite{spark}. 
All code was implemented in Python 3.6.

There are some notable differences between the GNU Parallel and Apache Spark implementations.
For instance, in the GNU Parallel implementation, all operations (i.e. reading, incrementing and writing)
are done within a single task. This task is applied to all the \bigbrain blocks using GNU Parallel,
processing a subset of them at a time,
depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
is achieved using the popular neuroimaging I/O library: \href{https://nipy.org/nibabel/}{NiBabel}. When provided with a filename, as is the case
for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memory map the data
using \href{https://numpy.org/}{NumPy}. This step will be referred to as ``load header''. It is only when the data is actually required
(i.e. during incrementation), that the data will be loaded into memory.

The Apache Spark implementation differs from that of GNU Parallel. Whereas GNU Parallel will
simply forks processes for each task and defers scheduling to the kernel, Spark is responsible for 
task scheduling decisions. Furthermore, in our Spark implementation, we opted to read the data using Spark's built-in
\texttt{BinaryFiles}, loading whole files, in binary format, into Spark partitions. As the
data would have been pre-loaded by Spark, ``loading header'' only measures the time to
convert the binary images into NiBabel objects, and ``incrementation'' consists solely
of the time it took to increment the data. Therefore, only the write time would be
comparable between the two implementations, as it is achieved using the same method.
Furthermore, the Spark implementation also differs in that read, increment and write
were separated into three map tasks, which can enable shuffling between the tasks should 
it be determined as necessary by the Spark scheduler.

@@

The BIDS App Example is a template example for creating a Brain Imaging Data Structure (BIDS)
compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
on all the anatomical (T1W) images of datasets containing numerous subjects. This step is 
referred to as Participant analysis within the application. An optional
step of the BIDS App example, referred to as Group analysis, computes the average brain
mask size of the entire dataset.

For our experiments, we used the entire CoRR dataset, made available on \href{https://www.datalad.org/}{DataLad} and
applied Participant analysis to it. The BIDS App Example was executed using a Singularity container stored on
Isilon. As in the BigBrain Incrementation, the experiment was parallelized 
using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode,
where Isilon and the local SSD were only evaluated in Memory Mode. Each experiment was repeated 3x.

Time in this application was obtained using Linux's \texttt{time()} application.
\texttt{Real + Sys} was used to measure CPU time,
whereas \texttt{User - (Real + Sys)} was used to measure
I/O time.
\section{Results}

\subsection{40~$\mu$m \bigbrain incrementation}
\subsubsection{25 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Makespan of the \bigbrain Incrementation application using 25 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. Three 
             repetitions were performed. *Local SSD did not complete the
         writing of the last few blocks (approximately 5-10) due to storage limitations.}\label{fig:makespan-25cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_25cpus.pdf}
    \captionsetup{belowskip=0pt}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. Three repetitions were performed. *Local SSD did not
    complete the writing of the last few blocks (approximately 5-10) due to storage limitations}\label{fig:stacked-25cpus}
\end{figure}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572380543-tmpfsAD_1it_25cpus_40bb-1.pdf}
    \caption{DRAM}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382482-optaneAD_1it_25cpus_40bb-1.pdf}
    \caption{Optane DCPMM}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382794-localAD_1it_25cpus_40bb-1.pdf}
    \caption{local SSD*}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572388118-isilonAD_1it_25cpus_40bb-2.pdf}
    \caption{Isilon}\label{fig:gantt25isilon}
\end{subfigure}
    \captionsetup{belowskip=-10pt}
    \caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes. *Some local SSD writes (approximately 5-10 writes) did not complete due to storage limitations.}\label{fig:gantt25}
\end{figure*}

@@

The App Direct mode Gantt charts (Figure~\ref{fig:gantt25}) also reflect what is observed in Figure~\ref{fig:stacked-25cpus}.
DRAM is significantly faster than all other storage, leading it to barely appear within the Gantt chart.
DRAM was measured to have an average of 17.2 parallel tasks throughout the execution. Optane DCPMM takes
significantly longer than DRAM and appears to spend a significant amount of time writing. The average
parallelism measured for Optane DCPMM was 24. Unlike Optane DCPMM, local SSD spends the vast majority of its time
reading, while spending a negligible amount of time writing, due to writeback to DRAM. Like Optane DCPMM, it averages 24 parallel tasks
throughout its execution. Unlike the other three storage, Isilon shows spacing between the loading of the header,
read and increment and write. Isilon was measured to have an average parallelism of 14. Overall, each storage device
execution displayed heterogeneous task durations.

\subsubsection{96 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_96cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Makespan of the \bigbrain Incrementation application using 96 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. Three 
             repetitions were performed. *Local SSD did not complete the
         writing of the last few blocks (approximately 5-10) due to storage limitations.}\label{fig:makespan-96cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_96cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. Three repetitions were performed. *Local SSD did not
    complete the writing of the last few blocks (approximately 5-10) due to storage limitations}\label{fig:stacked-96cpus}
\end{figure}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572306855-tmpfs_1it_96cpus_40bb-2.pdf}
    \caption{Memory mode}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572379747-optaneAD_1it_96cpus_40bb-1.pdf}
    \caption{App Direct mode}
\end{subfigure}
    \captionsetup{belowskip=-10pt}
\caption{Gantt charts for Optane DCPMM processing 125 blocks of the 40$\mu$m BigBrain using 96 processes}\label{fig:gantt96}
\end{figure*}

Despite having increased parallelism by a factor of 3.84, we see no visible reduction
in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 processes, DRAM and Optane DCPMM in Memory mode are the only devices that did better than the anticipated I/O, with a makespan duration of
5.7s and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
Optane DCPMM in App Direct mode and local SSD both performed worse than the anticipated I/O estimates, Isilon
performed as expected. Variance was found to be high on the local SSD.

@@

A further look at the Gantt charts between Optane DCPMM in Memory mode and App Direct mode (Figure~\ref{fig:gantt96})
shows that both read and write time of tasks are worse in App Direct mode as compared to Memory mode.
While some read tasks in App Direct mode appear to be of similar duration to the average read task time
of Memory mode, the vast majority of read tasks are significantly slower. Furthermore, no write tasks in App
Direct mode is capable of reaching the performance of equivalent tasks in Memory mode.

@@

Unlike the GNU Parallel executions, the Spark executions appear to be in better accordance with
the I/O estimates (Figure~\ref{fig:20mkspspark25}). Only Optane DCPMM in App Direct mode appears
to have significantly exceeded the estimates, by a factor of 2. Total write duration all appear
to be longer in the Spark implementation than with GNU Parallel (Figure~\ref{fig:20totalspark:25}).
Spark's writes are 1.4-1.6x longer with Optane DCPMM and 1.1x longer with Isilon. Due to differences in
how data is loaded in Spark, it is not possible to comment on how read times are affected. However, it
is possible to note that data conversion (binary string to NumPy array) takes almost twice as 
much time on Optane DCPMM than it does on Isilon. Furthermore, the act of incrementation can take
up to twice as much time on Isilon.

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_25cpus.pdf}
    \caption{Makespan}\label{fig:20mksp25}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
    \caption{Total read/increment/write breakdown}\label{fig:20total25}
\end{subfigure}
\captionsetup{belowskip=-10pt}
\caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
25 processes. Three repetitions were performed.}\label{fig:2025}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark_20bb_25cpus.pdf}
    \caption{Makespan}\label{fig:20mkspspark25}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_25cpus.pdf}
    \caption{Total convert/increment/write breadown}\label{fig:20totalspark:25}
\end{subfigure}
\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes. Three repetitions wereperformed}\label{fig:20spark25}
\end{figure*}
